Predicting Remaining Useful Life of NASA Aircraft Engines

### 2025-03-30 index.qmd
3/30 updates

summarymetrics <- summary(Minimum = min(time),
                          Median = median(time),
                          StandardDeviation = sd(time),
                          Maximum = max(time), 
                          distributions$time)

kable(tibble(summarymetrics), format = "markdown")

@@

```{r}
cox <- coxph(Surv(time, status) ~ os1 + os3 + sm1 + sm2 + sm5,
             data = train,
             id = id)

cox1 <- coxph(data = train, 
             Surv(time, status) ~ os1 + os3 + sm1 + sm2 + sm5 + sm6 + sm7 + sm11 + sm12 + sm13 + sm14 + sm19 + sm20 + sm21)

cox2 <- coxph(data = train, 
             Surv(time, status) ~ os1 + os3 + sm1 + sm2 + sm5)

cox3 <- coxph(data = train, 
             Surv(time, status) ~ os1 + os3 + sm1 + sm5 + sm6 + sm7 + sm13 + sm14 + sm19 + sm21)

cox4 <- coxph(data = train, 
             Surv(time, status) ~ sm5 + sm6 + sm7 + sm11 + sm12 + sm13 + sm14 + sm19 + sm20 + sm21)

cox5 <- coxph(data = train, 
             Surv(time, status) ~ os1 + os3 + sm1 + sm2 + sm5  + sm7 + sm11 + sm12 + sm13 + sm19 + sm20 )

cox6 <- coxph(data = train, 
             Surv(time, status) ~ os1  + sm1 + sm2 + sm5  + sm7 + sm11 + sm12 + sm13 + sm19  + sm21)

cox7 <- coxph(data = train, 
             Surv(time, status) ~ os1  + sm13 + sm14 + sm19 + sm20)

cox8 <- coxph(data = train, 
             Surv(time, status) ~ os1 + os3 + sm21)

cox9 <- coxph(data = train, 
             Surv(time, status) ~ os1 + os3 + sm2 + sm5 + sm6 + sm12 + sm13  + sm19 + sm20 + sm21)

cox10 <- coxph(data = train, 
             Surv(time, status) ~  os3 + sm1 + sm2 + sm5 + sm6  + sm11 + sm12 + sm13  + sm19 + sm20 + sm21)

@@

### Model Evaluation

@@

ggsurvplot(survfit, data = test)
```

@@

Create df with max time for each individual to use in predictions

```{r}
#predictiondata <- test %>%
  #group_by(id) %>%
  #slice_max(time) %>%
  #select(id, time, os1, os3, sm1, sm2, sm5, status)

#coef <- data.frame(Variable = c("Z1", "Z2", "Z3", "Z4", "Z5"),
           #CoefficientValue = cox1.1$coefficients)
#kable(coef)
```

Hazard Rate Calculation:
$h(t|Z) = h_0(t) * e^{B_1Z_1 + B_2Z_2 + B_3Z_3+B_4Z_4+B_5Z_5}$

```{r}
test$hazardrates <- predict(cox1.1,
        newdata = test,
        type = "risk")

@@

@@

```{r}
test <- test %>%
  arrange(id, time) %>%
  group_by(id) %>%
  mutate(cumulative_hazard = cumsum(hazardrates))
```

Predicting survival probability

```{r}
test$survprob <- predict(cox1.1, type = "survival", newdata = test)

@@

Probability Calculation: $S(t) = e^{-H(t)}$ $H(t)$ is the cumulative
hazard up to time $t$

Tim until event can be modeled in R by passing the `type = "lp"`
argument in the `stats::predict()` function to get a value for linear
predictor.

@@

A study was conducted to determine how...Modeling and Results

-   Explain your data preprocessing and cleaning steps. transformation
    applied to data to add event

-   Present your key findings in a clear and concise manner.

-   Use visuals to support your claims.

-   **Tell a story about what the data reveals.**

### Results

>>>>

summarymetrics <- data.frame(c("Minimum", "Median", "Mean", "Standard Deviation", "Maximum"),
                             c(min(distributions$time), 
                               median(distributions$time),
                               mean(distributions$time),
                               sd(distributions$time),
                               max(distributions$time)))

colnames(summarymetrics) <- c("Metric", "Value")
summarymetrics$Value <- round(summarymetrics$Value, digits = 2)
kable(summarymetrics)

@@

```{r}
cox1 <- coxph(data = train, id = id,
             Surv(time, status) ~ os1 + os3 + sm1 + sm2 + sm5 + sm6 + sm7 + sm11 + sm12 + sm13 + sm14 + sm19 + sm20 + sm21)

cox2 <- coxph(data = train, id = id,
             Surv(time, status) ~ os1 + os3 + sm1 + sm2 + sm5)

cox3 <- coxph(data = train, id = id,
             Surv(time, status) ~ os1 + os3 + sm1 + sm5 + sm6 + sm7 + sm13 + sm14 + sm19 + sm21)

cox4 <- coxph(data = train, id = id,
             Surv(time, status) ~ sm5 + sm6 + sm7 + sm11 + sm12 + sm13 + sm14 + sm19 + sm20 + sm21)

cox5 <- coxph(data = train, id = id,
             Surv(time, status) ~ os1 + os3 + sm1 + sm2 + sm5  + sm7 + sm11 + sm12 + sm13 + sm19 + sm20 )

cox6 <- coxph(data = train, id = id,
             Surv(time, status) ~ os1  + sm1 + sm2 + sm5  + sm7 + sm11 + sm12 + sm13 + sm19  + sm21)

cox7 <- coxph(data = train, id = id,
             Surv(time, status) ~ os1  + sm13 + sm14 + sm19 + sm20)

cox8 <- coxph(data = train, id = id,
             Surv(time, status) ~ os1 + os3 + sm21)

cox9 <- coxph(data = train, id = id,
             Surv(time, status) ~ os1 + os3 + sm2 + sm5 + sm6 + sm12 + sm13  + sm19 + sm20 + sm21)

cox10 <- coxph(data = train, id = id,
             Surv(time, status) ~  os3 + sm1 + sm2 + sm5 + sm6  + sm11 + sm12 + sm13  + sm19 + sm20 + sm21)

@@

@@

@@

The model created can be used to calculate several relevant metrics when predicting remaining time for the engines until expected failure. Predictions will be made on the largest value for time for each individual engine.

#### Model Coefficients

The CPH model created produced coefficients for each covariate indicating the quantitative effect on the resulting hazard rate. The coefficients are shown below.

```{r}
coef <- data.frame(CoefficientValue = cox1.1$coefficients)
kable(coef)
```

#### Hazard Rate Calculation

The coefficients above can be used to calculate the hazard rate at time $t$. The hazard rate will be calculated for each engine at the engines max time value.

Hazard Rate:
$h(t|Z) = h_0(t) * e^{B_1Z_1 + B_2Z_2 + B_3Z_3+B_4Z_4+B_5Z_5}$

The engines with the higher hazard rates are at a higher risk of experiencing failure.

```{r}
test$HazardRate <- predict(cox1.1,
        newdata = test,
        type = "risk")

predictiondata <- test %>%
  group_by(id) %>%
  slice_max(time) %>%
  mutate(CumulativeHazard = cumsum(HazardRate)) %>%
  select(id, time, sm2, sm5, sm7, sm11, sm12, sm20, sm21, status, HazardRate)

kable(head(predictiondata %>% select(id, time, sm2, sm5, sm7, sm11, sm12, sm20, sm21, HazardRate)))

@@

#### Cumulative Hazard Calculation

The cumulative hazard represents the total hazard accumulated for the individual over time and like the hazard rate, is another measure that can be used to estimate the risk for the individual at a point in time. The cumulative hazard for each individual is shown below. Higher values represent a higher chance of failure.

@@

```{r}
kable(head(predictiondata))

@@

#### Survival probability

While cumulative hazard can be used by itself to give an idea of an individual's risk, it can also be used to calculate a probability of survival.

Probability Calculation: $S(t) = e^{-H(t)}$ $H(t)$ is the cumulative
hazard up to time $t$.

```{r}
predictiondata$SurvivalProbability <- predict(cox1.1, type = "survival", newdata = predictiondata)
kable(head(predictiondata))

#### Predicting Time Until Failure

Time until event can be modeled in R by passing the `type = "lp"`
argument in the `stats::predict()` function to get a value for linear
predictor.

@@
